### 2024-11-23 project2/nnMetamodels.tex
new

\begin{algorithm}
\caption{Two-Stage Metamodeling Nested Simulation Procedure for Estimating CVaR}
\begin{algorithmic}[1] \label{alg2:twoStageProcedure}
    \STATE \textbf{Train a neural network metamodel using simulation data:}
    \begin{enumerate} [label=\alph*., itemsep=0pt, parsep=0pt, topsep=0pt]
        \item Use a fraction of the total simulation budget to run Steps 1, 2, and 3 in the standard procedure, i.e., Algorithm~\ref{alg2:standardProcedure}, with the same number of outer scenarios, $M$, but a much smaller number of inner replications, $N' \ll N$, in each scenario. 
        Obtain $M$ simulated samples (feature-label pairs), $(\bS^{(i)}, \Lhat_i)$, $i=1,\ldots,M$. 
        Note that $N'$ may be much smaller than $N$, so $\Lhat_i$ are expected to have larger variance.
        \item Use the simulated data, $(\bS^{(i)}, \Lhat_i)$, $i=1,\ldots,M$ to train a supervised learning model. 
        It is standard practice in supervised learning research to transform the stock prices $\bS^{(i)}$ to log returns $\mathbf{X}^{(i)}$,
        $$
        \mathbf{X}^{(i)} = \left(\ln\left(\frac{S_{1}^{(i)}}{S_{0}^{(i)}}\right), \ln\left(\frac{S_{2}^{(i)}}{S_{1}^{(i)}}\right), \ldots, \ln\left(\frac{S_{T}^{(i)}}{S_{T-1}^{(i)}}\right)\right),
        $$
        and use $\mathbf{X}^{(i)}$ as features instead of the stock prices $\bS^{(i)}$.
        Refer to the trained model as a metamodel and denote it by $\Lhat^{PD}(\mathbf{X})$. Denote the predicted losses for the outer scenarios by $\Lhat^{PD}_i = \Lhat^{PD}(\mathbf{X}^{(i)})$, $i=1,\ldots,M$.
        \item Sort the predicted losses $\Lhat^{PD}_{(1)}\leq \Lhat^{PD}_{(2)} \leq \cdots\leq \Lhat^{PD}_{(M)}$ to identify a predicted tail scenario set, $\widehat{\tail}_{m}^{PD}$, associated with the largest predicted losses. The number of predicted tail scenarios, $m$, is a user's choice.
    \end{enumerate}
    \STATE \textbf{Concentrate simulation on predicted tail scenarios:}
    \begin{enumerate} [label=\alph*., itemsep=0pt, parsep=0pt, topsep=0pt]
        \item Run Steps 2 and 3 of Algorithm~\ref{alg2:standardProcedure} with the same number of inner replications, $N$, but only on the predicted tail scenarios, i.e., scenarios in $\widehat{\tail}_{m}^{PD}$. Denote the standard procedure's estimated losses and sorted losses by $\Lhat^{ML}_i$ and $\Lhat^{ML}_{(i)}$, respectively, $i=1,\ldots,m$.
        \item Estimate the $\alpha$-CVaR of $L$ by 
        $$\widehat{\CVaR}^{ML}_\alpha = \frac{1}{(1-\alpha)M} \sum_{i=\alpha M + 1}^{M}\Lhat_{(i)}^{ML} = \frac{1}{(1-\alpha)M} \sum_{i \in \widehat{\tail}_{(1-\alpha )M}^{ML}}\Lhat_{i}^{ML},$$ 
        where $\widehat{\tail}_{k}^{ML}$ denotes a predicted tail scenario set associated with the largest $k$ estimated losses.
    \end{enumerate}
\end{algorithmic}
\end{algorithm}

>>>>

\begin{algorithm}
\caption{Two-Stage Metamodeling Nested Simulation Procedure for Estimating CVaR}
\begin{algorithmic}[1] \label{alg2:twoStageProcedure}
    \STATE \textbf{Train a neural network metamodel using simulation data:}
    \begin{enumerate} [label=\alph*., itemsep=0pt, parsep=0pt, topsep=0pt]
        \item Use a fraction of the total simulation budget to run Steps 1, 2, and 3 in the standard procedure, i.e., Algorithm~\ref{alg2:standardProcedure}, with the same number of outer scenarios, $M$, but a much smaller number of inner replications, $N' \ll N$, in each scenario. 
        Obtain $M$ simulated samples (feature-label pairs), $(\bS^{(i)}, \Lhat_i)$, $i=1,\ldots,M$. 
        Note that $N'$ may be much smaller than $N$, so $\Lhat_i$ are expected to have larger variance.
        \item Use the simulated data, $(\bS^{(i)}, \Lhat_i)$, $i=1,\ldots,M$ to train a supervised learning model. 
        It is standard practice in supervised learning research to transform the stock prices $\bS^{(i)}$ to returns $\mathbf{X}^{(i)}$,
        \begin{equation} \label{eq2:return}
        \mathbf{X}^{(i)} = \left(\frac{S_1^{(i)} - S_0}{S_0}, \frac{S_2^{(i)} - S_1^{(i)}}{S_1^{(i)}}, \ldots, \frac{S_T^{(i)} - S_{T-1}^{(i)}}{S_{T-1}^{(i)}}\right)
        \end{equation}
        and use $\mathbf{X}^{(i)}$ as features instead of the stock prices $\bS^{(i)}$.
        Refer to the trained model as a metamodel and denote it by $\Lhat^{PD}(\mathbf{X})$. Denote the predicted losses for the outer scenarios by $\Lhat^{PD}_i = \Lhat^{PD}(\mathbf{X}^{(i)})$, $i=1,\ldots,M$.
        \item Sort the predicted losses $\Lhat^{PD}_{(1)}\leq \Lhat^{PD}_{(2)} \leq \cdots\leq \Lhat^{PD}_{(M)}$ to identify a predicted tail scenario set, $\widehat{\tail}_{m}^{PD}$, associated with the largest predicted losses. The number of predicted tail scenarios, $m$, is a user's choice.
    \end{enumerate}
    \STATE \textbf{Concentrate simulation on predicted tail scenarios:}
    \begin{enumerate} [label=\alph*., itemsep=0pt, parsep=0pt, topsep=0pt]
        \item Run Steps 2 and 3 of Algorithm~\ref{alg2:standardProcedure} with the same number of inner replications, $N$, but only on the predicted tail scenarios, i.e., scenarios in $\widehat{\tail}_{m}^{PD}$. Denote the standard procedure's estimated losses and sorted losses by $\Lhat^{ML}_i$ and $\Lhat^{ML}_{(i)}$, respectively, $i=1,\ldots,m$.
        \item Estimate the $\alpha$-CVaR of $L$ by 
        $$\widehat{\CVaR}^{ML}_\alpha = \frac{1}{(1-\alpha)M} \sum_{i=\alpha M + 1}^{M}\Lhat_{(i)}^{ML} = \frac{1}{(1-\alpha)M} \sum_{i \in \widehat{\tail}_{(1-\alpha )M}^{ML}}\Lhat_{i}^{ML},$$ 
        where $\widehat{\tail}_{k}^{ML}$ denotes a predicted tail scenario set associated with the largest $k$ estimated losses.
    \end{enumerate}
\end{algorithmic}
\end{algorithm}

### 2024-11-23 project3/draft.tex
new

LSTM networks are well-suited for modeling sequential data due to their ability to capture long-term dependencies.
For the application of VA risk management using metamodel-based nested simulation, LSTM networks approximate the inner simulation, i.e., the mapping from market scenarios to the scenario-wise contract losses.
In Chapter~\ref{chap:project2}, we treat metamodeling as a supervised learning problem and demonstrated that LSTM networks can effectively model this complex relationships with extensive training on a large dataset of VA simulations.
The total computational cost originates from two sources: running the standard nested simulation procedure to generate the training data and training the LSTM network on the generate dataset.
Given a new of VA contract, the above process needs to be repeated to adapt the LSTM metamodel to the new contract.
For a DNN with a large number of parameters, retraining the LSTM network from scratch can be computationally expensive.
TL offers an alternative approach to keep the previous knowledge as a foundation, thus accelerates the adaptation of the LSTM metamodel to new conditions.
Computational savings come in two forms: 
\begin{equation*}
    \item   less training time is required to fine-tune the pre-trained model on the new dataset, and
    \item   fewer data points are needed to achieve a good LSTM metamodel, which is particularly beneficial when the standard nested simulation procedure is costly.
\end{equation*}

@@

A typical TL framework for supervised learning consists of the following:
\begin{itemize}
    \item   \textbf{source domain}: $\mathcal{D}_{\text{So}} = \{\mathcal{X}_{\text{So}}, F_{\text{So}}(X)\}$
    \item   \textbf{source task}: $\mathcal{T}_{\text{So}} = \{\mathcal{Y}_{\text{So}}, f_{\text{So}}(\cdot)\}$
    \item   \textbf{target domain}: $\mathcal{D}_{\text{Ta}} = \{\mathcal{X}_{\text{Ta}}, F_{\text{Ta}}(x)\}$
    \item   \textbf{target task}: $\mathcal{T}_{\text{Ta}} = \{\mathcal{Y}_{\text{Ta}}, f_{\text{Ta}}(\cdot)\}$
\end{itemize}

where $\mathcal{X}_{\text{So}}$ and $\mathcal{X}_{\text{Ta}}$ include input features derived from the outer simulation of two different VA contracts or market conditions, and $F_{\text{So}}(X)$ and $F_{\text{Ta}}(x)$ are the marginal probability distributions of the source and target domains, respectively.
In our numerical experiments, the dataset is generated by running the standard nested simulation procedure in Algorithm~\ref{alg2:standardProcedure} for a large number of VA contracts.
The input features $X$ are the risk factors, and the output labels $L$ are the contract losses at each time step.
We keep the number of features to be $240$ for both source and target domains as in Section~\ref{sec2:numerical}.
The predictive function $f_{\text{So}}$ and $f_{\text{Ta}}$ are trained to estimate outputs $L_{\text{So}}$ and $L_{\text{Ta}}$ such as contract losses in inner simulations under the source and target tasks, respectively.
TL seeks to improve the learning of the target predictive function $f_{\text{Ta}}(\cdot)$ in $\mathcal{D}_{\text{Ta}}$ by leveraging the previous training on $\mathcal{D}_{\text{So}}$ and $f_{\text{So}}(\cdot)$, particularly when $\mathcal{D}_{\text{So}} \neq \mathcal{D}_{\text{Ta}}$ or $\mathcal{T}_{\text{So}} \neq \mathcal{T}_{\text{Ta}}$.
In Chapter~\ref{chap:project2}, $\mathcal{D}_{\text{So}} = \mathcal{D}_{\text{Ta}}$ and $\mathcal{T}_{\text{So}} = \mathcal{T}_{\text{Ta}}$.

In our TL setting, both the source and target tasks involve learning a mapping from the risk factors to the VA contract losses.
For the source task, we have a dataset $\mathcal{D}_{\text{So}} = { (X_{\text{So}}^{(i)}, L_{\text{So}}^{(i)}) }_{i=1}^{M_{\text{So}}}$, where $M_{\text{So}}$ is the number of training samples, i.e., the number of outer simulation paths used to generate the dataset with a standard nested simulation procedure.
$X_{\text{So}}^{(i)} \in \mathcal{X}_{\text{So}}$ and $L_{\text{So}}^{(i)} \in \mathcal{Y}_{\text{So}}$.

The ultimate objective is to learn a metamodel $f_{\text{Ta}}$ that predicts the VA contract losses $L_{\text{So}}$ from the scenarios $X_{\text{Ta}}$ in the form of a financial time series.
TL starts by training a metamodel $f_{\text{So}}$ on the source domain $\mathcal{D}_{\text{So}}$.
$f_{\text{So}}$ is approximated by an LSTM network $f_{\text{So}}(\cdot ; \theta_{\text{So}})$, and the parameters of the LSTM network, $\theta_{\text{So}}$ are learned by minimizing a MSE loss function on the source domain.
Then the pre-trained parameters $\theta_{\text{So}}$ to inform the learning of $f_{\text{Ta}}(\cdot ; \theta_{\text{Ta}})$ on the target domain $\mathcal{D}_{\text{Ta}}$.
The training on the target domain should encourage similarity between $\theta_{\text{So}}$ and $\theta_{\text{Ta}}$ to facilitate the transfer of knowledge.

>>>>

LSTM networks are well-suited for modeling sequential data due to their ability to capture long-term dependencies.
For the application of VA risk management using metamodel-based nested simulation, LSTM networks approximate the inner simulation, i.e., the mapping from market scenarios to the scenario-wise contract losses.
In Chapter~\ref{chap:project2}, we treat metamodeling as a supervised learning problem and demonstrated that LSTM networks can effectively model this complex relationships with extensive training on a large dataset of VA simulations.
The total computational cost originates from two sources: running the standard nested simulation procedure to generate the training data and training the LSTM network on the generate dataset.
Given a new of VA contract, the above process needs to be repeated to adapt the LSTM metamodel to the new contract.
For a DNN with a large number of parameters, retraining the LSTM network from scratch can be computationally expensive.
TL offers an alternative approach to keep the previous knowledge as a foundation, thus accelerates the adaptation of the LSTM metamodel to new conditions.
Computational savings come in two forms: 
\begin{enumerate}
    \item   less training time is required to fine-tune the pre-trained model on the new dataset, and
    \item   fewer data points are needed to achieve a good LSTM metamodel, which is particularly beneficial when the standard nested simulation procedure is costly.
\end{enumerate}

@@

A typical TL framework for supervised learning consists of the following:
\begin{itemize}
    \item   \textbf{source domain}: $\mathcal{D}_{\text{So}} = \{\mathcal{X}_{\text{So}}, F_{\text{So}}(\mathbf{X})\}$
    \item   \textbf{source task}: $\mathcal{T}_{\text{So}} = \{\mathcal{Y}_{\text{So}}, f_{\text{So}}(\cdot)\}$
    \item   \textbf{target domain}: $\mathcal{D}_{\text{Ta}} = \{\mathcal{X}_{\text{Ta}}, F_{\text{Ta}}(\cdot)\}$
    \item   \textbf{target task}: $\mathcal{T}_{\text{Ta}} = \{\mathcal{Y}_{\text{Ta}}, f_{\text{Ta}}(\cdot)\}$
\end{itemize}

where $\mathcal{X}_{\text{So}}$ and $\mathcal{X}_{\text{Ta}}$ include input features derived from the outer simulation of two different VA contracts or market conditions, and $F_{\text{So}}$ and $F_{\text{Ta}}$ are the marginal probability distributions of the source and target domains, respectively.
In our numerical experiments, the dataset is generated by running the standard nested simulation procedure in Algorithm~\ref{alg2:standardProcedure} for a large number of VA contracts.+-
The input features $\mathbf{X}$ are the return vectors from Equation~\ref{eq2:return}, and the output labels $L$ are the contract losses at each time step.
The stock price vector $\mathbf{S}$ is not crucial for the discussion of this chapter and is omitted since it can be immediately recovered from $\mathbf{X}$ and the initial stock price $S_0$.
We keep the number of features to be $240$ for both source and target domains as in Section~\ref{sec2:numerical}.
The predictive function $f_{\text{So}}$ and $f_{\text{Ta}}$ are trained to estimate outputs $L_{\text{So}}$ and $L_{\text{Ta}}$ such as contract losses in inner simulations under the source and target tasks, respectively.
TL seeks to improve the learning of the target predictive function $f_{\text{Ta}}(\cdot)$ in $\mathcal{D}_{\text{Ta}}$ by leveraging the previous training on $\mathcal{D}_{\text{So}}$ and $f_{\text{So}}(\cdot)$, particularly when $\mathcal{D}_{\text{So}} \neq \mathcal{D}_{\text{Ta}}$ or $\mathcal{T}_{\text{So}} \neq \mathcal{T}_{\text{Ta}}$.
In Chapter~\ref{chap:project2}, $\mathcal{D}_{\text{So}} = \mathcal{D}_{\text{Ta}}$ and $\mathcal{T}_{\text{So}} = \mathcal{T}_{\text{Ta}}$.

In our TL setting, both the source and target tasks involve learning a mapping from the risk factors to the VA contract losses.
For the source task, we have a dataset $\mathcal{D}_{\text{So}} = { (X_{\text{So}}^{(i)}, L_{\text{So}}^{(i)}) }_{i=1}^{M_{\text{So}}}$, where $M_{\text{So}}$ is the number of training samples, i.e., the number of outer simulation paths used to generate the dataset with a standard nested simulation procedure.
$X_{\text{So}}^{(i)} \in \mathcal{X}_{\text{So}}$ and $L_{\text{So}}^{(i)} \in \mathcal{Y}_{\text{So}}$.
The ultimate objective is to learn a metamodel $f_{\text{Ta}}$ that predicts the VA contract losses $L_{\text{So}}$ from the scenarios $X_{\text{Ta}}$ in the form of a financial time series.
TL starts by training a metamodel $f_{\text{So}}$ on the source domain $\mathcal{D}_{\text{So}}$.
$f_{\text{So}}$ is approximated by an LSTM network $f_{\text{So}}(\cdot ; \theta_{\text{So}})$, and the parameters of the LSTM network, $\theta_{\text{So}}$ are learned by minimizing a MSE loss function on the source domain.
Then the pre-trained parameters $\theta_{\text{So}}$ to inform the learning of $f_{\text{Ta}}(\cdot ; \theta_{\text{Ta}})$ on the target domain $\mathcal{D}_{\text{Ta}}$.
The training on the target domain should encourage similarity between $\theta_{\text{So}}$ and $\theta_{\text{Ta}}$ to facilitate the transfer of knowledge.
